---
title: "Data integration"
author: "Zhiyuan Hu"
date: "07/01/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(SingleCellExperiment)) #1.4.1
suppressPackageStartupMessages(library(scater)) #1.10.1
suppressPackageStartupMessages(library(edgeR)) #3.24.3
suppressPackageStartupMessages(library(limma)) #3.38.3
suppressPackageStartupMessages(library(dplyr)) #0.7.8
# devtools::install_version(package = 'Seurat', version = package_version('3.0.2'))
library(Seurat)#3.0.2
library(scales) # 1.0.0
library(reshape2) #1.4.3
# devtools::install_github('chris-mcginnis-ucsf/DoubletFinder')
library(DoubletFinder) #2.0.1
set.seed(16475)
# library(reticulate)
# py_install("umap-learn")
```



```{r read data, include=FALSE}
sceset <- readRDS("../../man_analysis2_20180629/rds/20190214_allFT_Clincluster_12clusters_sceset_withUMAP.rds")
sceset <- calculateQCMetrics(sceset)
sc16 <- readRDS("../../20190216scRNA-seq16/clean_data/sceset_sc16.rds")
sc16 <- calculateQCMetrics(sc16)
scb1 <- readRDS("../../20190825scRNA-seq_benign/clean_data/sceset_allBenign.rds")
scb1 <- calculateQCMetrics(scb1)

counts <- cbind(counts(sceset), counts(sc16)[match(rownames(sceset), rownames(sc16)),] ,counts(scb1)[match(rownames(sceset), rownames(scb1)),])
coldata_scb1 <- colData(scb1)
coldata_scb1$source <- "Fresh"
coldata_scb1$final.clusters <- NA
coldata_sceset <- colData(sceset)
coldata_sceset$FACS <- "FTE"
coldata_sc16 <- colData(sc16)
coldata_sc16$source <- "Fresh"
coldata_sc16$final.clusters <- NA
coldata_sceset <- coldata_sceset[, colnames(coldata_sceset) %in% colnames(coldata_scb1)]
coldata_sc16 <- coldata_sc16[, colnames(coldata_sc16) %in% colnames(coldata_sceset)]

colnames(coldata_sceset)[!colnames(coldata_sceset) %in% colnames(coldata_sc16)]

coldata <- rbind(coldata_sceset, coldata_sc16[match(colnames(coldata_sceset), colnames(coldata_sc16))], coldata_scb1[match(colnames(coldata_sceset), colnames(coldata_scb1))])

rowdata <- rowData(sceset)

sceset <- SingleCellExperiment(assays=list(counts = counts))
rowData(sceset) <- rowdata
colData(sceset) <- coldata
rm(rowdata)
logcounts(sceset) <- log2(calculateCPM(sceset) + 1)

sceset$type <- "Cancer"
sceset$type[grep(x = sceset$batch, pattern = "scb")] <- "Benign"
rm(scb1)
rm(sc16)
sceset <- sceset[,sceset$source == "Fresh"]

sceset$FACS[sceset$FACS == "CD45-EPCAM+"] <- "FTE"
sceset$FACS[sceset$FACS == "EPCAM+CD45-"] <- "FTE"
sceset$FACS[sceset$FACS == "CD45-EPCAM-"] <- "STROMAL"
sceset$FACS <- toupper(sceset$FACS)

# UMAP
# too very long time to run
set.seed(16475)
sceset <- runUMAP(sceset, ncomponents = 2, 
  exprs_values = "logcounts", scale_features = TRUE)

set.seed(16475)
sceset <- runTSNE(sceset, ncomponents = 2, 
  exprs_values = "logcounts", scale_features = TRUE)

rm(coldata, coldata_sc16, coldata_scb1, coldata_sceset, counts)
gc()
```


```{r add secretory annotations, include=T}
## Add secretory annotations
secretory <- readRDS("../../man_analysis2_20180629/rds/20190120Fresh_secretory_9clusters_clincluster.rds")

secretory$cell_type[secretory$clincluster_final == "C10"] <- "Cell cycle (C9)"
secretory$cell_type[secretory$clincluster_final == "C3"] <- "Differentiated (C3)"
secretory$cell_type[secretory$clincluster_final == "C4"] <- "KRT17 (C4)"
secretory$cell_type[secretory$clincluster_final == "C6"] <- "Stress (C6)"
secretory$cell_type[secretory$clincluster_final == "C8"] <- "EMT (C7)"
secretory$cell_type[secretory$clincluster_final == "C9"] <- "Immune (C8)"
secretory$cell_type[secretory$clincluster_fina %in% c("C1","C2","C5")] <- "Queiscent"
secretory$secretory_subtype <- secretory$cell_type
cancer_secretory <- secretory
```


```{r doubletFinder, include=FALSE, message=FALSE, eval=FALSE}
sceset$secretory_subtype <- secretory$cell_type[match(colnames(sceset), colnames(secretory))]

seurat <- as.Seurat(sceset)
seurat <- NormalizeData(seurat)
seurat <- ScaleData(object = seurat)
seurat <- FindVariableFeatures(seurat, selection.method = "vst", nfeatures = 2000)
seurat <- RunPCA(seurat)
seurat <- RunUMAP(seurat, dims = 1:15, seed.use = 12345 )
seurat <- RunTSNE(seurat, dims = 1:15, seed.use = 12345 )

seurat <- FindNeighbors(seurat, dims = 1:15)
seurat <- FindClusters(seurat, resolution = 0.6)
# DimPlot(seurat, reduction = "tsne", label = T)

seurat$disease <- NA
seurat$disease[seurat$Patient2 %in% c(11543, 11545, 11553)] <- "ovarian"
seurat$disease[seurat$Patient2 %in% c(15066, 15091, 15072)] <- "endometrial"
seurat$disease[seurat$Patient2 %in% c(33572, 33778, 34350, 34659,35773 )] <- "benign"

sweep.res.list_kidney <- paramSweep_v3(seurat, PCs = 1:15, sct = FALSE)
sweep.stats_kidney <- summarizeSweep(sweep.res.list_kidney, GT = FALSE)
bcmvn_kidney <- find.pK(sweep.stats_kidney)

## Homotypic Doublet Proportion Estimate 
annotations <- seurat$seurat_clusters ## Clustering results
homotypic.prop <- modelHomotypic(annotations)           ## ex: annotations <- seu_kidney@meta.data$ClusteringResults
nExp_poi <- round(0.01*length(seurat@active.ident))  ## 1% doublet rate
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop)) ## 

## Run DoubletFinder with varying classification stringencies
seurat <- doubletFinder_v3(seurat, PCs = 1:10, pN = 0.25, pK = 0.005, nExp = nExp_poi, reuse.pANN = FALSE)
seurat <- doubletFinder_v3(seurat, PCs = 1:10, pN = 0.25, pK = 0.005, nExp = nExp_poi.adj, reuse.pANN = "pANN_0.25_0.005_43")

seurat@meta.data$DF_hi.lo <- seurat@meta.data$DF.classifications_0.25_0.005_43
seurat@meta.data$DF_hi.lo[which(seurat@meta.data$DF_hi.lo == "Doublet" & seurat@meta.data$DF.classifications_0.25_0.005_36 == "Singlet")] <- "Doublet_lo"
seurat@meta.data$DF_hi.lo[which(seurat@meta.data$DF_hi.lo == "Doublet")] <- "Doublet_hi"

# DimPlot(seurat, reduction = "tsne", group.by="DF_hi.lo", plot.order=c("Doublet_hi","Doublet_lo","Singlet"), colors.use=c("black","gold","red"))

colData(sceset)$DF_hi.lo <- seurat$DF_hi.lo
# all(colnames(seurat)==colnames(sceset))

benign_ft <- seurat[,seurat$DF_hi.lo == "Singlet" & seurat$type=="Benign"]
# saveRDS(benign_ft, "rds/20191030benignFT_1857cells_SeuratObj.rds")

```

```{r read-rds}
## read in preprocessed data
benign_ft <- readRDS("../../scFT-paper_rds/20191030benignFT_1857cells_SeuratObj.rds")
```

```{r umap-by-patients, fig.height=4, fig.width=10}
p1 <-  DimPlot(benign_ft, reduction = "TSNE", group.by = "Patient2", pt = 0.4) + 
    ggtitle(paste("t-SNE plot \n Benign FT (n = ", ncol(benign_ft), ")", sep = ""))
p2 <- DimPlot(benign_ft, reduction = "UMAP", group.by = "Patient2", pt = 0.4) + 
    ggtitle(paste("UMAP plot \n Benign FT (n = ", ncol(benign_ft), ")", sep = ""))

cowplot::plot_grid(p1,p2)
# ggsave("../revision_plots/SI/FigS3_benignFT_UMAPtSNE_colByPatient.png", width = 10, height = 4)
```


```{r, fig.height=6, fig.width=12}
FeaturePlot(benign_ft, reduction = "UMAP",features = c("PTPRC","COL1A1","EPCAM","KRT7","PAX8","CAPS"), ncol  = 3, pt.size = 0.4)
# ggsave("../revision_plots/SI/FigS3_benignFT_UMAPcolourByMarkers.png", width = 14, height = 8)
```

```{r}
DimPlot(benign_ft, reduction = "UMAP",  pt = 0.4, label = T) 
```


## Filter out non-secretory cells

```{r}
table(seurat$seurat_clusters[seurat$type == "Benign" & seurat$DF_hi.lo == "Singlet"])
#  0   1   2   3   4   5   6   7   8   9  10  11  12 
# 198 922   9 509  38   1  33   6  79  19  40   0   3 
```


```{r filter-stroma-lymphocytes-cilitaed}
# sum(benign_ft$seurat_clusters %in% c(8,10:12) | benign_ft$FACS == "CD45+EPCAM-") # 122 leukocytes
secretory <- benign_ft[,!benign_ft$seurat_clusters %in% c(8,10:12) & benign_ft$FACS != "CD45+EPCAM-"] # remove CD45+ cells

# sum(secretory$FACS == "STROMAL") # 11 stroma cells
secretory <- secretory[, !secretory$FACS %in% c( "STROMAL")]

dim(secretory)
```

1724 FTE cells in total

```{r Filter by expression}
# sum(secretory$seurat_clusters[secretory$type== "Benign"] %in% c(3, 7)) # 515 ciliated cells
secretory <- secretory[, !secretory$seurat_clusters %in% c(3, 7)] # remove cilicated cells
# 

cpm <- log2(calculateCPM(secretory@assays$RNA@counts) + 1)
secretory <- secretory[, (cpm["KRT7",] > 0.5 | cpm["PAX8",] > 0.5) &
                         cpm["EPCAM",] > 0.5  &
                           cpm["PTPRC",] == 0 #&
                           # cpm["CCDC17",] < 1
                       ]
rm(cpm)
table(secretory$Patient2)
# 33572 33778 34350 34659 35773 
#   210    53   130   356    16 
secretory_new <- secretory
# saveRDS(secretory, "rds/20191030benign_765secretory_seuratObj.rds")
```


```{r}
cancer_secretory <- as.Seurat(cancer_secretory)
cancer_secretory$type <- "Cancer"
secretory <- merge(secretory_new, cancer_secretory)

secretory$secretory_subtype[is.na(secretory$secretory_subtype) & secretory$type == "Cancer"] <- "Unclassified"

# saveRDS(secretory, "rds/seurat_secretory_merged20191007.rds")
```

## Merged data (unintegrated)

```{r}
secretory <- NormalizeData(secretory)
secretory <- ScaleData(object = secretory)
secretory <- FindVariableFeatures(secretory, selection.method = "vst", nfeatures = 2000)
secretory <- RunPCA(secretory)
ElbowPlot(secretory, 40)

secretory <- RunUMAP(secretory, dims = 1:14, seed.use = 123345)
secretory <- RunTSNE(secretory, dims = 1:14, seed.use = 123345)

secretory <- FindNeighbors(secretory, dims = 1:14, seed.use = 123345)
secretory <- FindClusters(secretory, resolution = 0.6)
```

Batch effects existed between the two batches.

```{r umaps-show-batch-effects,fig.height=3, fig.width=8}
p1 <- DimPlot(secretory, reduction = "umap", label = T)
p2 <- DimPlot(secretory, reduction = "umap", group.by = "type")
cowplot::plot_grid(p1,p2)
```


## Data integration (standard workflow)

This section integrated different batches

```{r data-integration-seurat, fig.width=8, fig.height=3}
seu_list <- list()

seu_list[[1]] <- secretory[,secretory$type == "Cancer"]
seu_list[[1]] <- seu_list[[1]][, seu_list[[1]]$secretory_subtype != "Immune (C8)" & seu_list[[1]]$secretory_subtype != "Unclassified"]

seu_list[[2]] <- secretory[,secretory$Patient2 %in% c("34350","33778","35773","33572")]
seu_list[[3]] <- secretory[,secretory$Patient2 == "34659"]

for(itor in 1:3){
    seu_list[[itor]] <- NormalizeData(seu_list[[itor]], verbose = FALSE)
    seu_list[[itor]] <- FindVariableFeatures(seu_list[[itor]], 
                                             selection.method = "vst", 
                                             nfeatures = 1000, verbose = FALSE)
}

n_dim <- 30
ftsc.anchors <- FindIntegrationAnchors(object.list = seu_list, 
                                       dims = 1:n_dim,
                                       k.anchor = 5) 

ftsc.integrated <- IntegrateData(anchorset = ftsc.anchors, dims = 1:n_dim)

DefaultAssay(ftsc.integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
ftsc.integrated <- ScaleData(ftsc.integrated, verbose = FALSE)
ftsc.integrated <- RunPCA(ftsc.integrated, npcs = n_dim, verbose = FALSE)
ftsc.integrated <- RunUMAP(ftsc.integrated, reduction = "pca", dims = 1:n_dim)
p1 <- DimPlot(ftsc.integrated, reduction = "umap", group.by = "type", pt.size = 0.5)
p2 <- DimPlot(ftsc.integrated, 
              reduction = "umap", 
              group.by = "secretory_subtype", pt.size = 0.5)
cowplot::plot_grid(p1,p2, rel_widths = c(1,1.1))
```



```{r, fig.height=3, fig.width=8}
p1 <- DimPlot(ftsc.integrated[,ftsc.integrated$type == "Cancer" & !is.na(ftsc.integrated$secretory_subtype)], reduction = "umap", group.by = "secretory_subtype") + ggtitle("Cancer")
p2 <- DimPlot(ftsc.integrated[,ftsc.integrated$type == "Benign"],group.by = "type",reduction = "umap") + ggtitle("Benign")
cowplot::plot_grid(p1,p2)
```


#### Clustering


```{r}
ftsc.integrated <- FindNeighbors(ftsc.integrated, dims = 1:n_dim, seed.use = 12345, k.param = 7)
ftsc.integrated <- FindClusters(ftsc.integrated, resolution = 0.6)

DimPlot(ftsc.integrated, label = T, reduction = "umap")
```


```{r}
table(ftsc.integrated$secretory_subtype[ftsc.integrated$type == "Cancer"], 
      ftsc.integrated$seurat_clusters[ftsc.integrated$type == "Cancer"])
```


```{r}
table(ftsc.integrated$type, ftsc.integrated$seurat_clusters)
```

```{r}
ftsc.integrated$secretory_subtype_benign <- NA
ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Benign" & ftsc.integrated$seurat_clusters %in% c(4)] <- "KRT17"

ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Benign" & ftsc.integrated$seurat_clusters %in% c(10)] <- "Cell cycle"

ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Benign" & ftsc.integrated$seurat_clusters %in% c(9)] <- "EMT"

ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Benign" & ftsc.integrated$seurat_clusters %in% c(4)] <- "Differentiated"


ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Cancer" & ftsc.integrated$secretory_subtype == "KRT17 (C4)"] <- "KRT17"

ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Cancer" & ftsc.integrated$secretory_subtype == "Cell cycle (C9)"] <- "Cell cycle"

ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Cancer" & ftsc.integrated$secretory_subtype == "EMT (C7)"] <- "EMT"

ftsc.integrated$secretory_subtype_benign[ftsc.integrated$type == "Cancer" & ftsc.integrated$secretory_subtype == "Differentiated (C3)"] <- "Differentiated"
```


```{r}
table(ftsc.integrated$secretory_subtype_benign, ftsc.integrated$type)
```


```{r}
DimPlot(ftsc.integrated, reduction = "umap", label = T, pt.size = 0.5)

# ggsave("../revision_plots/SI/integrated_secretory_umap_seuratResults.png", width = 5, height = 4)
```


```{r, fig.width=11, fig.height=3}
p1 <- DimPlot(ftsc.integrated[,!is.na(ftsc.integrated$secretory_subtype_benign)], reduction = "umap", label = F,group.by = "secretory_subtype_benign", pt.size = 0.5) + ggtitle("Normal FT secretory cells \n from all patients") + theme(legend.position = "none")
p2 <- DimPlot(ftsc.integrated[,!is.na(ftsc.integrated$secretory_subtype_benign) & ftsc.integrated$type == "Cancer"], reduction = "umap", label = F, group.by = "secretory_subtype_benign", pt.size = 0.5) + ggtitle("Normal FT secretory cells \n from cancer patients")+ theme(legend.position = "none")
p3 <- DimPlot(ftsc.integrated[,!is.na(ftsc.integrated$secretory_subtype_benign) & ftsc.integrated$type == "Benign"], reduction = "umap", label = F, group.by = "secretory_subtype_benign", pt.size = 0.5) + ggtitle("Normal FT secretory cells \n from benign patients")
p4 <- DimPlot(ftsc.integrated[,!is.na(ftsc.integrated$secretory_subtype_benign) & ftsc.integrated$type == "Benign"], reduction = "umap", label = F, group.by = "Patient2", pt.size = 0.5) + ggtitle("Normal FT secretory cells \n from benign patients (colored by patients)")
cowplot::plot_grid(p1,p2,p3, p4, nrow = 1, rel_widths = c(1,1,1.25,1.15))

# ggsave("../revision_plots/SI/integrated_secretory_umap.png", width = 17, height = 4)
```



```{r}
integrated.markers <- FindAllMarkers(ftsc.integrated, only.pos = T, assay = "RNA", random.seed = 12345)
```


#### Scoring the cells from benign samples

```{r}
new_markers <- readRDS("../R_deconvolution/rds/20190213new_markers.rds")
ftsc.integrated$KRT17_score <- colSums(ftsc.integrated@assays$RNA@data[new_markers$C4,])
ftsc.integrated$CC_score <- colSums(ftsc.integrated@assays$RNA@data[new_markers$C10,])
ftsc.integrated$EMT_score <- colSums(ftsc.integrated@assays$RNA@data[new_markers$EMT,])
ftsc.integrated$Diff_score <- colSums(ftsc.integrated@assays$RNA@data[new_markers$C3,])
```

```{r, fig.width=12, fig.height=3}
ftsc.sceset <- as.SingleCellExperiment(ftsc.integrated)

df.plot <- cbind(ftsc.sceset@reducedDims$UMAP, 
                 ftsc.sceset@colData[,c("KRT17_score", "CC_score", "EMT_score", "Diff_score", "type", "secretory_subtype_benign")])
df.plot <- as.data.frame(df.plot)
# df.plot <- df.plot[!is.na(df.plot$secretory_subtype_benign),]
df.plot1 <- df.plot[df.plot$type == "Benign",]
df.plot2 <- df.plot[df.plot$type == "Cancer",]

scaleScore <- function(x, max.limit = 2.5) {
    x <- scale(x, center = T, scale = T)
    x[x > max.limit] <- max.limit
    return(x)
}

df.plot1$KRT17_score <- scaleScore(df.plot1$KRT17_score)
df.plot1$CC_score <- scaleScore(df.plot1$CC_score)
df.plot1$Diff_score <- scaleScore(df.plot1$Diff_score)
df.plot1$EMT_score <- scaleScore(df.plot1$EMT_score)


df.plot2$KRT17_score <- scaleScore(df.plot2$KRT17_score)
df.plot2$CC_score <- scaleScore(df.plot2$CC_score)
df.plot2$Diff_score <- scaleScore(df.plot2$Diff_score)
df.plot2$EMT_score <- scaleScore(df.plot2$EMT_score)


plotScoreUmap <- function(dat, column){
        ggplot(data=dat, aes_string(x="UMAP_1", y="UMAP_2", fill = column)) + 
        geom_point(alpha = 0.5, pch = 21, col = "grey80") +
        theme_classic2() +
        scale_fill_gradient2(midpoint=0,
                              low="grey90", mid="white",high="red", space ="Lab" ) +
        theme(plot.title = element_text(size=14, face="bold", hjust = 0.5))
}

plist <- list()
plist[[1]] <- plotScoreUmap(dat = df.plot1, column = "KRT17_score") + ggtitle("Benign (KRT17 score)")
plist[[2]] <- plotScoreUmap(dat = df.plot1, column = "CC_score") + ggtitle("Benign (cell cycle score)")
plist[[3]] <- plotScoreUmap(dat = df.plot1, column = "Diff_score") + ggtitle("Benign (differentiated score)")
plist[[4]] <- plotScoreUmap(dat = df.plot1, column = "EMT_score") + ggtitle("Benign (EMT score)")
plist[[5]] <- plotScoreUmap(dat = df.plot2, column = "KRT17_score") + ggtitle("Cancer (KRT17 score)")
plist[[6]] <- plotScoreUmap(dat = df.plot2, column = "CC_score") + ggtitle("Cancer (cell cycle score)")
plist[[7]] <- plotScoreUmap(dat = df.plot2, column = "Diff_score") + ggtitle("Cancer (differentiated score)")
plist[[8]] <- plotScoreUmap(dat = df.plot2, column = "EMT_score") + ggtitle("Cancer (EMT score)")

cowplot::plot_grid(plotlist = plist, ncol = 4)

# ggsave("../revision_plots/SI/benign_cancer_subtype_score_plots.png", height = 8, width = 18)
# ggsave("../revision_plots/SI/benign_cancer_subtype_score_plots.tiff", height = 8, width = 18)

# ggsave("../revision_plots/SI/benign_cancer_subtype_score_plots.pdf", height = 8, width = 18)
```


## Save rds

```{r}
# saveRDS(ftsc.integrated, "rds/ftsc.integrated_20191013.rds")
```


```{r}
# p1 <- DimPlot(secretory.integrated, reduction = "umap", label = T)
# p2 <- DimPlot(secretory.integrated, reduction = "umap", group.by = "disease")
# p3 <- DimPlot(secretory.integrated[,!is.na(secretory.integrated$secretory_subtype)], reduction = "umap", group.by = "secretory_subtype",cols = my_colours[c(1:5,7,9)])
# 
# cowplot::plot_grid(p1,p2,p3, nrow = 1, rel_widths = c(1,1.2,1.4))
# ggsave("plots/20190917/umap_integrated.png", width = 18, height = 5)
```

$C3
[1] "LTBP4"    "SLC25A25" "LAMC2"    "DHCR24"   "PLK3"     "LRG1"     "LDLR"    

$C4
 [1] "SPP1"     "IL1B"     "IL1RN"    "KRT23"    "ALDH3B2"  "SUSD2"    "DEFB1"    "HLA-DQA2" "CYP4B1"   "PIGR"    

$EMT
 [1] "SPARC"    "SERPINF1" "DCN"      "SFRP4"    "CRISPLD2" "TIMP3"    "CNN1"     "MYH11"    "MFAP4"    "ENG"      "EFEMP1"  
[12] "RGS16"   

$C10
 [1] "FEN1"   "NUSAP1" "UBE2C"  "ZWINT"  "PRC1"   "ASF1B"  "MCM4"   "GINS2"  "CENPM"  "MCM2"   "TK1"    "MCM6"   "SMC4"  
[14] "MLF1IP" "MAD2L1"

$ciliated
[1] "TEKT1"    "TUBA4B"   "C20orf85" "CAPSL"    "LRRC46"   "EFCAB1"   "C6orf118" "CCDC78"  



## Technical

```{r}
sessionInfo()
```

R version 3.5.2 (2018-12-20)
Platform: x86_64-apple-darwin15.6.0 (64-bit)
Running under: macOS Mojave 10.14.4

Matrix products: default
BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/3.5/Resources/lib/libRlapack.dylib

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] class_7.3-15                DoubletFinder_2.0.1         reshape2_1.4.3              scales_1.0.0               
 [5] Seurat_3.0.2                dplyr_0.7.8                 edgeR_3.24.3                limma_3.38.3               
 [9] scater_1.10.1               ggplot2_3.2.0.9000          SingleCellExperiment_1.4.1  SummarizedExperiment_1.12.0
[13] DelayedArray_0.8.0          BiocParallel_1.16.5         matrixStats_0.54.0          Biobase_2.42.0             
[17] GenomicRanges_1.34.0        GenomeInfoDb_1.18.1         IRanges_2.16.0              S4Vectors_0.20.1           
[21] BiocGenerics_0.28.0        

loaded via a namespace (and not attached):
  [1] backports_1.1.3          plyr_1.8.4               igraph_1.2.3             lazyeval_0.2.1          
  [5] splines_3.5.2            listenv_0.7.0            usethis_1.4.0            digest_0.6.18           
  [9] htmltools_0.3.6          viridis_0.5.1            gdata_2.18.0             magrittr_1.5            
 [13] memoise_1.1.0            cluster_2.0.7-1          ROCR_1.0-7               remotes_2.0.2           
 [17] globals_0.12.4           RcppParallel_4.4.3       R.utils_2.7.0            prettyunits_1.0.2       
 [21] colorspace_1.4-0         ggrepel_0.8.0            xfun_0.4                 callr_3.1.1             
 [25] crayon_1.3.4             RCurl_1.95-4.11          jsonlite_1.6             bindr_0.1.1             
 [29] survival_2.43-3          zoo_1.8-4                ape_5.2                  glue_1.3.0              
 [33] gtable_0.2.0             zlibbioc_1.28.0          XVector_0.22.0           leiden_0.3.1            
 [37] pkgbuild_1.0.2           Rhdf5lib_1.4.2           future.apply_1.3.0       HDF5Array_1.10.1        
 [41] bibtex_0.4.2             Rcpp_1.0.0               metap_1.1                viridisLite_0.3.0       
 [45] reticulate_1.10          rsvd_1.0.2               SDMTools_1.1-221         tsne_0.1-3              
 [49] htmlwidgets_1.3          httr_1.4.0               gplots_3.0.1.1           RColorBrewer_1.1-2      
 [53] ica_1.0-2                pkgconfig_2.0.2          R.methodsS3_1.7.1        uwot_0.1.3              
 [57] locfit_1.5-9.1           tidyselect_0.2.5         labeling_0.3             rlang_0.4.0             
 [61] munsell_0.5.0            tools_3.5.2              cli_1.0.1                devtools_2.0.1          
 [65] ggridges_0.5.1           stringr_1.4.0            yaml_2.2.0               npsurv_0.4-0            
 [69] fs_1.2.6                 processx_3.2.1           knitr_1.21               fitdistrplus_1.0-14     
 [73] caTools_1.17.1.1         purrr_0.3.0              RANN_2.6.1               bindrcpp_0.2.2          
 [77] pbapply_1.4-0            future_1.14.0            nlme_3.1-137             R.oo_1.22.0             
 [81] compiler_3.5.2           rstudioapi_0.9.0         curl_3.3                 beeswarm_0.2.3          
 [85] plotly_4.8.0             png_0.1-7                lsei_1.2-0               tibble_2.0.1            
 [89] stringi_1.2.4            ps_1.3.0                 desc_1.2.0               RSpectra_0.13-1         
 [93] lattice_0.20-38          Matrix_1.2-15            pillar_1.3.1             Rdpack_0.10-1           
 [97] lmtest_0.9-36            RcppAnnoy_0.0.11         data.table_1.12.0        cowplot_0.9.4           
[101] bitops_1.0-6             irlba_2.3.3              gbRd_0.4-11              R6_2.3.0                
[105] KernSmooth_2.23-15       gridExtra_2.3            vipor_0.4.5              sessioninfo_1.1.1       
[109] codetools_0.2-16         MASS_7.3-51.1            gtools_3.8.1             assertthat_0.2.0        
[113] pkgload_1.0.2            rhdf5_2.26.2             rprojroot_1.3-2          withr_2.1.2             
[117] sctransform_0.2.0        GenomeInfoDbData_1.2.0   grid_3.5.2               tidyr_0.8.2             
[121] DelayedMatrixStats_1.4.0 Rtsne_0.15               ggbeeswarm_

