---
title: "Visulisation of secretory clusters"
author: "Zhiyuan Hu"
date: "12/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE)
library(SingleCellExperiment) #1.4.1
library(scater)#1.10.1
library(edgeR) #3.24.3
library(limma) #3.38.3
library(dplyr) #0.7.8
library(scales)
library(reshape2)
library(DoubletFinder) # 2.0.1
# devtools::install_version(package = 'Seurat', version = package_version('2.3.4'))
library(Seurat) # 2.3.4
```

## Summary

In this part, we are going to visualise the secretory cells after clustering.

It will include the figures we presented in our manuscript.

```{r read clustering results}
sceset <- readRDS("../../scFT-paper_rds/20190214_allFT_Clincluster_12clusters_sceset_withUMAP.rds")

secretory <- readRDS("../../scFT-paper_rds/20190120Fresh_secretory_9clusters_clincluster.rds")
markers2 <- read.csv("../../scFT-paper_rds/20190120Clincluster_fresh_secretory_9clusters_markers.csv", as.is = T)
```

## Visualisation

#### Scatter plot of secretory clusters

```{r scran-normalisation}
secretory <- scran::computeSumFactors(secretory)
secretory <- scater::normalize(secretory)
```


```{r runTSNE-runUMAP}
set.seed(123456)
secretory <- runTSNE(object = secretory, ncomponents = 2, 
                     feature_set = rownames(secretory)[rowData(secretory)$high.var],
                  exprs_values = "normcounts",
                  perplexity = min(50, floor(ncol(secretory)/5)))
secretory <- scater::runUMAP(object = secretory, ncomponents = 2, use_dimred = "PCA", n_dimred = 15,exprs_values = "normcounts")
```

```{r assign-human-readable-lable-to-clusters}
secretory$cell_type <- NA
secretory$cell_type[secretory$clincluster_final == "C10"] <- "Cell cycle (C9)"
secretory$cell_type[secretory$clincluster_final == "C3"] <- "Differentiated (C3)"
secretory$cell_type[secretory$clincluster_final == "C4"] <- "KRT17 (C4)"
secretory$cell_type[secretory$clincluster_final == "C6"] <- "Stress (C6)"
secretory$cell_type[secretory$clincluster_final == "C8"] <- "EMT (C7)"
secretory$cell_type[secretory$clincluster_final == "C9"] <- "Immune (C8)"

secretory$Cell_subtype <- secretory$cell_type
```


```{r plot-TSNE-by-celltypes-or-by-patients, fig.width=8, fig.height=3}
p1 <- plotTSNE(secretory[,secretory$clincluster_final %in% c("C3","C4","C8","C10","C9")], colour_by = "Cell_subtype") +
    # theme(legend.position = "top" )  + 
    xlab("TSNE_1") + ylab("TSNE_2")

p2 <- plotTSNE(secretory[,secretory$clincluster_final %in% c("C3","C4","C8","C10","C9")], colour_by = "Patient2") +
    # theme(legend.position = "top" )  + 
    xlab("TSNE_1") + ylab("TSNE_2")
cowplot::plot_grid(p1, p2, rel_widths = c(1.1, 1))
# ggsave("plots/20190508_TSNE_secretory.png", width=5, height = 3.5)
```

```{r plot-umap-by-celltypes, fig.width=4, fig.height=3}
plotUMAP(secretory[,secretory$clincluster_final %in% c("C3","C4","C8","C10","C9")], colour_by = "Cell_subtype") +
    # theme(legend.position = "top" )  + 
    xlab("UMAP_1") + ylab("UMAP_2")
# ggsave("../../revision_analysis_20190827/revision_plots/SI/secretory_UMAP_colSubtype20191007.png", width=5, height = 3.5)
```

```{r plot-tsne-by-patient, fig.width=4, fig.height=3}
plotTSNE(secretory[,secretory$clincluster_final %in% c("C3","C4","C8","C10","C9")], colour_by = "Patient2") +
    # theme(legend.position = "top" )  + 
    xlab("TSNE_1") + ylab("TSNE_2")
# ggsave("../../revision_analysis_20190827/revision_plots/SI/secretory_TSNE_colPatient20191007.png", width=5, height = 3.5)
```

```{r plot-tsne-by-celltypes, fig.width=4, fig.height=3}
plotTSNE(secretory[,secretory$clincluster_final %in% c("C3","C4","C8","C10","C9")], colour_by = "Cell_subtype") +
    # theme(legend.position = "top" )  + 
    xlab("TSNE_1") + ylab("TSNE_2")
# ggsave("../../revision_analysis_20190827/revision_plots/main/secretory_TSNE_colSubtype20191007.png", width=5, height = 3.5)
```

```{r}
plotUMAP(secretory[,secretory$clincluster_final %in% c("C3","C4","C8","C10")], colour_by  = "Cell_subtype")
```


#### Scatter plots of fresh FTE cells

Plot UMAP of all fresh cells

```{r plot-UMAP-of-all-fresh-cells}
# ciliated.markers <- read.csv("../tables/20180725allFT_clincluster_markers.csv", as.is = T, row.names = 1)
# ciliated.markers <- ciliated.markers$gene[ciliated.markers$logFC > 5 & ciliated.markers$cluster == 11]
fresh <- sceset[,sceset$source == "Fresh"]

# hvg only in clude markers of the four secretory cells and ciliated markers
# feature.set <- c(markers2$gene[markers2$cluster %in% c("C10","C3","C4","C8")], ciliated.markers)
# set.seed(14567)
# fresh <- runUMAP(fresh, ncomponents = 2, feature_set = rownames(fresh) %in% feature.set,
#    exprs_values = "logcounts", scale_features = TRUE)

fresh$cell_type <- NA
fresh$cell_type[colnames(fresh) %in% colnames(sceset)[sceset$final.clusters == 11]] <- "Ciliated"
fresh$cell_type[colnames(fresh) %in% colnames(secretory)] <- "Secretory"

table(fresh$cell_type)
```

```{r plotUMAP-fresh-byPatients, fig.width=4, fig.height=3}
plotUMAP(fresh, colour_by = "Patient") + xlab("UMAP_1") + ylab("UMAP_2")
# p2 <- plotUMAP(fresh, 
#          colour_by = "Patient2") + xlab("UMAP_1") + ylab("UMAP_2")
# cowplot::plot_grid(p1,p2)
# ggsave("exprloratory_plots/20190508UMAP_fresh_only.png", height = 3, width = 6)
```


```{r plotUMAP-fresh-byCAPSs, fig.width=8, fig.height=3}
p1 <- plotUMAP(fresh, colour_by = "CAPS") + xlab("UMAP_1") + ylab("UMAP_2")
p2 <- plotUMAP(fresh, colour_by = "KRT7") + xlab("UMAP_1") + ylab("UMAP_2")
cowplot::plot_grid(p1,p2)
```


#### Heatmap

```{r heatmap-secretory, fig.height=5, fig.width=8}
top10 <- markers2 %>% group_by(cluster) %>% top_n(15, logFC)
top10 <- top10[!top10$cluster %in% c("C1","C2","C5") ,]
top10 <- top10[top10$logFC > 1.8, ]

top10 <- top10[order(top10$cluster, decreasing = F),]

plot.data <- logcounts(secretory)[c(top10$gene), order(secretory$clincluster_final, decreasing = F)]

colanno <- data.frame (colData(secretory)[,c("clincluster_final","Patient2")])
colnames(colanno)[1] <- "clusters"
colanno$clusters <- factor(colanno$clusters)

rownames(colanno) <- colnames(secretory)
colanno <- colanno[order(colanno$clusters, decreasing = F),]
colanno$clusters <- factor(colanno$clusters, levels = unique(colanno$clusters))
plot.data <- plot.data[,match(rownames(colanno), colnames(plot.data))] 

plot.data <- t(scale(t(plot.data), center = T, scale = T))
# range(plot.data )
plot.data <- Seurat::MinMax(plot.data, min = -2.5, max = 2.5)

plot.data<- as.data.frame(x = t(x = plot.data))
plot.data$cell <- rownames(x = plot.data)

cells.ident <- secretory$clincluster_final
names(x = cells.ident) <- secretory$Sample

colnames(x = plot.data) <- make.unique(names = colnames(x = plot.data))
plot.data %>% melt(id.vars = "cell") -> plot.data
names(x = plot.data)[names(x = plot.data) == 'variable'] <- 'gene'
names(x = plot.data)[names(x = plot.data) == 'value'] <- 'expression'
plot.data$ident <- cells.ident[plot.data$cell]

plot.data$gene <- with(
    data = plot.data,
    expr = factor(x = gene, levels = rev(x = unique(x = plot.data$gene)))
    )
plot.data$cell <- with(
    data = plot.data,
    expr = factor(x = cell, levels = unique(x = colnames(secretory)))
    )

plot.data$ident <- factor(plot.data$ident, levels = c("C1","C2","C10","C3","C4","C5","C6","C8","C9"))
# http://colorbrewer2.org/#type=sequential&scheme=Blues&n=6

heatmap <- ggplot( data = plot.data, mapping = aes(x = cell, y = gene, fill = expression)) + 
    geom_tile() +
    scale_fill_gradient2( # colour
        low = muted("steelblue4"), mid = "white",
        high = muted("firebrick2")
        # low = "purple", mid = "black",
        # high = "yellow"
        ) +
    scale_y_discrete(position = "right", labels = rev(top10$gene)) +
    theme(
      axis.line = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      strip.text.x = element_text(size = 12),
      axis.text.y = element_text(size=6, face="italic"),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      strip.background = element_blank()
    )

heatmap <- heatmap +
      facet_grid(
        facets = ~ident,
        drop = TRUE,
        space = "free",
        scales = "free",
        switch = 'x'
      ) +
      scale_x_discrete(expand = c(0, 0), drop = TRUE)  
      
panel.spacing <- unit(x = 0.15, units = 'lines')
heatmap <- heatmap +
      theme(panel.spacing = panel.spacing)

heatmap

# ggsave(filename = "../manuscript_plots/secretory subtypes/Figure_secretory_heatmap_9clusters_yellowPurple.pdf", height = 18, width = 24, units = "cm" )
# ggsave(filename = "../manuscript_plots/secretory subtypes/Figure_secretory_heatmap_9clusters.png",dpi = 300, height = 18, width = 24, units = "cm" )
```

<!-- "The other protein of great interest, MCM2, is one of six proteins (MCM 2–7) that comprise the MCM helicase complex involved in the regulation of the S phase of the cell cycle in eukaryotes via the initiation of DNA replication (Kang et al, 2014). High MCM2 expression is reportedly a poor prognostic factor in ovarian cancer (Gakiopoulou et al, 2007). Furthermore, another member of the MCM2 family, MCM7, is reportedly a particularly poor prognostic factor in ovarian HGSC (Ota et al, 2011). In the present study, MCM2 was found to be differentially expressed in tubal and peritoneal HGSCs in addition to ovarian and endometrial HGSCs. "  (https://www.nature.com/articles/bjc201627) -->


```{r colour-bar-for-heatmap-by-clusters, fig.width=8, fig.height=2}
## Add colour bar
plot.data$patient <- secretory$Patient2[match(plot.data$cell, colnames(secretory))]
ggplot(plot.data, aes(x = cell, y = 1, fill = ident)) + geom_tile() + theme_void() +
      facet_grid(
        facets = ~ident,
        drop = TRUE,
        space = "free",
        scales = "free",
        switch = 'x'
      ) +
      theme(panel.spacing = panel.spacing)

```

```{r colour-bar-for-heatmap-by-patients, fig.width=8, fig.height=2}
ggplot(plot.data, aes(x = cell, y = 1, fill = patient)) + geom_tile() + theme_void() + #theme_classic() +
      facet_grid(
        facets = ~ident,
        drop = TRUE,
        space = "free",
        scales = "free",
        switch = 'x'
      ) +
      theme(panel.spacing = panel.spacing#, 
            # axis.text.x = element_text(size = 3, angle = 90)
            ) 
# ggsave("../../revision_analysis_20190827/revision_plots/main/Fig2B_bar.pdf", width = 12, height = 3)
```


#### Violin plots of marker genes

```{r violin-markers-cellcycle, fig.width=8, fig.height=6}
secretory$clincluster_final  <- as.factor(secretory$clincluster_final)

share.7a3.1 <- c( "MKI67","PCNA","CDK4","CDC45",
             "FANCD2","FANCI","MSH2","MSH6",
             "RUVBL2","RUVBL1","HMGB2","SMC1A")

plotExpression(secretory[,secretory$clincluster_final %in% c("C10","C3","C4","C6","C8")], 
               share.7a3.1,
               x = "clincluster_final", exprs_values = "logcounts", ncol = 4, xlab = "") + 
    scale_x_discrete(labels = list("C10"="CC","C3"="C3","C4"="KRT17","C8"="ECM")) + 
    theme(strip.text = element_text(face = "italic", size = 9))
          # axis.text.x = element_text(angle = 25, vjust = 0.5, hjust = 0)  )

# ggsave("../manuscript_plots/secretory subtypes/marker_cell_cycle_violin.pdf", height = 8, width = 12, units = "cm", dpi = 300)
```

Cluster 4 is characterized by upregulation of major histocompatibility complex (MHC) Class II genes (e.g. HLA-DQA1, HLA-DPA1 and HLA-DPB1), cytokeratins (KRT17, KRT5 and KRT23) and aldehyde dehydrogenases (e.g. ALDH1A1 and ALDH3B2) 

```{r violin-markers-krt17cluster, fig.width=8, fig.height=6}
secretory <- normalize(secretory)
assay(secretory, "normcounts") <- removeBatchEffect(assay(secretory, "normcounts"), covariates = secretory$total_features)
assay(secretory, "normcounts")[assay(secretory, "normcounts") <0] <- 0

# sseC4_marker <- c("HLA-DQA1", "HLA-DPA1", "HLA-DPB1",
#                "KRT17", "KRT23",
#                "ALDH1A1", "ALDH3B2","CDKN1A")
sseC4_marker <- c("HLA-DQA1", "HLA-DPA1",# "HLA-DPB1",
               "KRT17",  "KRT23",
               "ALDH1A1", "ALDH3B2")
plotExpression(secretory[,secretory$clincluster_final %in% c("C10","C3","C4","C8")], 
               sseC4_marker,
                x = "clincluster_final", exprs_values = "logcounts",
                ncol = 2, xlab = "") + ylab("Scaled expression") + 
    theme(strip.text = element_text(size = 10, face = "italic"))

# ggsave("plots/SuppFig3_marker_C4_20190508.png", width = 4, height = 4, dpi = 300)
```

```{r violin-markers-EMTclusters, fig.width=8, fig.height=6}
secretory$type <- NA
secretory$type[secretory$clincluster_final == "C10"] <- "Cell cycle"
secretory$type[secretory$clincluster_final == "C3"] <- "Differentiated"
secretory$type[secretory$clincluster_final == "C4"] <- "KRT17"
secretory$type[secretory$clincluster_final == "C8"] <- "EMT"
secretory$type <- factor(secretory$type, levels = c("Cell cycle", "Differentiated", "KRT17","EMT"))
plotExpression(secretory[,secretory$clincluster_final %in% c("C10","C3","C4","C8")], 
               c("EPCAM","KRT7","SPARC","RGS16"),
                x = "type", exprs_values = "logcounts",
                ncol = 2, xlab = "") + 
    theme(strip.text = element_text(size = 10, face = "italic"),
          axis.text.x = element_text(angle = 30, vjust = 0.7))
# ggsave("../../revision_analysis_20190827/revision_plots/main/Fig3a_marker_EMT_Epcam_KRT7.pdf", width = 6, height = 4, dpi = 300)
# ggsave("plots/SuppFig3_marker_EMT_Epcam_KRT7.png", width = 6, height = 4, dpi = 300)
```


```{r plotPhenoData-total_feature-by-clusters, fig.width=5, fig.height=3}
plotPhenoData(secretory, x = "clincluster_final", y = "total_features" )  + geom_hline(yintercept = 7500, col  = "grey50") + geom_hline(yintercept = 1200, col = "grey50") + xlab("Clusters")
# ggsave("plots/SuppFig3_marker_EMT_Epcam_KRT7.png", width = 6, height = 4, dpi = 300)
# ggsave("plots/SuppFig3_total_feautres.png", width = 5, height = 4)
```


## Technical

```{r tech-sessionInfo}
sessionInfo()
```

